Clustering time series based on probability distributions across temporal granularities

### 2021-11-18 main.Rmd
application rewording again

<!--Figure \ref{fig:opt-cluster-tsne-jsd} The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}. The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases. In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters in (a) and (b) respectively, and helps to characterize each cluster. In the quantile plots the line represents the median, and the region shows the area between the $25^{th}$ and $75^{th}$ percentiles. The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5. This additional division makes a clearer clustering, because it resolves the heterogeneity in $moy$ creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage. B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive $hod$ patterns but are both heterogeneous in $moy$ and $wnwd$. B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on $hod$ and $wnwd$. This clustering almost agrees with the clusters visible in the t-SNE plot.

@@

Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values $hod$, $moy$, and $wnwd$ of the $24$ customers, colored by $3$ (a) and $5$ (b) cluster solution as suggested by Figure \ref{fig:opt-cluster-tsne-jsd} through a parallel coordinate plot. The variables (wpd's for different granularities) are 
standardized prior to clustering using WPD. In the display, the variables are sorted according to their separation across groups. This means that $wnwd$ is the most important variable in distinguishing the groups, followed by $hod$ and $moy$ for both (a) and (b).
Group P-1 and P-2 correspond to Q-1 and Q-2 respectively. Cluster P-3 splits into Q-3, Q-4 and Q-5. Customers 16 and 18 are characterized by unusual high values of $wpd$ on $wnwd$ compared to the rest of the customers and hence form a group. This could again be verified from Figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their $wnwd$ behavior. They are represented by P-1. P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$. This can also be verified from Table \ref{tab:summary-table-wpd} where $hod$ has very low median values of $wpd$ over the members in the cluster, implying at least half of these customers do not have an interesting $hod$ pattern. P-3 behaves exactly opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$. These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at \ref{fig:hod-ind-group-png}.
For $5$ cluster solution, this groups gets split into Q3-Q5 characterizing different relative significance of $moy$ and $wnwd$. For example, Q-4 and Q-5 have almost no pattern across $moy$, but Q-3 has a $moy$ pattern and thus it is reasonable to split them. The patterns could be different, but they are significant. Q-4 and Q-5 are separated because of their significance across $hod$.

>>>>

<!--Figure \ref{fig:opt-cluster-tsne-jsd} The $sindex$ is plotted against the number of clusters for clustering based on JS-based (JS-NQT) distances in Figure \ref{fig:opt-cluster-tsne-jsd}. The $sindex$ falls from around $0.56$ when $k=3$ to $0.54$ when $k$ increases. In our study, we use $k=4, 5$.--> Figure \ref{fig:groups-4and5} displays the summarized distributions across $4$ and $5$ clusters from JS-NQT clustering in (a) and (b) respectively, and helps to characterize each cluster. In the quantile plots the line represents the median, and the region shows the area between the $25^{th}$ and $75^{th}$ percentiles. The only difference between the four and five cluster solution is that A-4 divides further into B-4 and B-5. This additional division makes a clearer clustering, because it resolves the heterogeneity in $moy$ creating a group (B-5, customers 1-3) which has a winter peak in usage, and a group (B-4, customers 16-20) which has a start of the year peak in usage. B-2 (customers 4-9) and B-1 (customers 21-24) have distinctive $hod$ patterns but are both heterogeneous in $moy$ and $wnwd$. B-3 (customers 10-15) has peak usage at the end of the year, but is heterogeneous on $hod$ and $wnwd$. This clustering almost agrees with the clusters visible in the t-SNE plot.

@@

Figure \ref{fig:summary-plot-wpd} shows the $wpd$ values of the $24$ customers over $hod$, $moy$, and $wnwd$ , colored by $3$ (a) and $5$ (b) clusters from WPD clustering using a parallel coordinate plot. The variables (wpd's for different granularities) are 
standardized prior to clustering using WPD. In the display, the variables are sorted according to their separation across groups. This means that $wnwd$ is the most important variable in distinguishing the groups, followed by $hod$ and $moy$ for both (a) and (b).
Groups P-1 and P-2 correspond to Q-1 and Q-2 respectively. Cluster P-3 splits into Q-3, Q-4 and Q-5. Customers 16 and 18 are characterized by unusual high values of $wpd$ on $wnwd$ compared to the rest of the customers and hence form a group. This could again be verified from Figure \ref{fig:hod-ind-group-png}, where these were the only two customers with a difference in their $wnwd$ behavior. They are represented by P-1. P-2 has lower $wpd$ for $hod$ than $moy$ and $wnwd$. P-3 behaves opposite to P-2 with higher $wpd$ for $hod$ compared to $moy$ and $wnwd$. These are the customers who have some significant pattern across $hod$ and this can again be validated by looking at Figure \ref{fig:hod-ind-group-png}. For a $5$ cluster solution, this group gets split into Q-3, Q-4 and Q-5 characterizing different relative significance of $moy$ and $wnwd$. For example, Q-4 and Q-5 have almost no pattern across $moy$, but Q-3 has a $moy$ pattern and thus it is reasonable to split them. The patterns could be different, but they are significant. Q-4 and Q-5 are separated because of their different significance of $hod$. All of these can also be verified from Table \ref{tab:summary-table-wpd} which shows the cluster summaries with members and median values of $wpd$ for the three variables.

<!-- $hod$ has very low median values of $wpd$ for the members in the cluster, implying at least half of these customers do not have an interesting $hod$ pattern. -->
